Power and Pitfalls of Generic Smart Contracts

### 2022-02-21 cluster_computing/010_introduction.tex
figures

In Solidity's bytecode,
non-primitive values are referenced through a very general
\<address> type. For instance, a Solidity method
\<child(Person p, uint256 n) returns Person> actually compiles
into \<child(address p, uint256 n) returns address>, losing most
type information~\cite{CrafaPZ19}.
Since, at run time, it is the bytecode that gets executed,
everything can be passed for \<p>, not just a \<Person> instance.
The compiler cannot even enforce strong typing
by generating defensive type instance checks and casts, because
values are unboxed in Ethereum: they have no attached
type information at run time,
they are just numerical \emph{addresses}.
It follows that inside the \<child> method, an eventual call to a \<Person>'s method
on \<p> might actually execute any arbitrary code, if \<p> is not a \<Person>.
In other words, Solidity is not strongly typed.
Consequently, it is highly discouraged, in Solidity, to call methods on parameters passed
to another method, such as on \<p> passed to \<child>, since an attacker can pass crafted
objects for \<p>, with arbitrary implementations for their methods,
which can result in the unexpected execution of
dangerous code. This actually happened in the case of the infamous DAO hack~\cite{dao16}, that
costed millions of dollars.

>>>>

In Solidity's bytecode,
non-primitive values are referenced through a very general
\<address> type. For instance, a Solidity method
\<child(Person p, uint256 n) returns Person> actually compiles
into \<child(address p, uint256 n) returns address>, losing most
type information~\cite{CrafaPZ19}.
Since, at run time, it is the bytecode that gets executed,
everything can be passed for \<p>, not just a \<Person> instance, as illustrated in Fig.\,\ref{figure.solidity_problem}.
The compiler cannot even enforce strong typing
by generating defensive type instance checks and casts, because
values are unboxed in Ethereum: they have no attached
type information at run time,
they are just numerical \emph{addresses}.
It follows that inside the \<child> method, an eventual call to a \<Person>'s method
on \<p> might actually execute any arbitrary code, if \<p> is not a \<Person>.
In other words, Solidity is not strongly typed.
Consequently, it is highly discouraged, in Solidity, to call methods on parameters passed
to another method, such as on \<p> passed to \<child>, since an attacker can pass crafted
objects for \<p>, with arbitrary implementations for their methods,
which can result in the unexpected execution of
dangerous code. This actually happened in the case of the infamous DAO hack~\cite{dao16}, that
costed millions of dollars.

\begin{figure}[ht]
\centering
\includegraphics[width=.9\linewidth]{figures/solidity_problem}
\caption{Example of possible problem that can occur in Solidity due to the absence of a strong typing mechanism.}
\label{figure.solidity_problem}
\end{figure}

### 2022-02-21 cluster_computing/015_generics_in_java.tex
figures

In order to understand the mechanism of erasure, consider for instance
the interface \<SharedEntity> in Fig.~\ref{fig:shared_entity} and its method \<accept>.
The functionality of \<SharedEntity> will be discussed later (Sec.~\ref{sec:shared_entities}).
Here, we only consider how its generic type parameters get compiled.
Namely, \<SharedEntity> uses two generic type parameters \<S> and \<O>, that must be provided
whenever a client creates a concrete implementation of the interface.
Such generic parameters have an upper bound: \<S> can only be a subtype of \<PayableContract>,
while \<O> can only be a subtype of {\codesize\texttt{Offer<S>}}. If we check the bytecode
generated for \<SharedEntity>, we will see that \<accept> is declared, in bytecode, as
\<void accept(BigInteger amount, PayableContract buyer, Offer offer)>, that is,
the two type variables \<S> and \<O> have been \emph{erased} and replaced with their
respective upper bound.

>>>>

In order to understand the mechanism of erasure, consider for instance
the interface \<SharedEntity> in Fig.~\ref{fig:shared_entity} and its method \<accept>.
The functionality of \<SharedEntity> will be discussed later (Sec.~\ref{sec:shared_entities}).
Here, we only consider how its generic type parameters get compiled.
Namely, \<SharedEntity> uses two generic type parameters \<S> and \<O>, that must be provided
whenever a client creates a concrete implementation of the interface.
Such generic parameters have an upper bound: \<S> can only be a subtype of \<PayableContract>,
while \<O> can only be a subtype of {\codesize\texttt{Offer<S>}}. If we check the bytecode
generated for \<SharedEntity>, we will see that \<accept> is declared, in bytecode, as
\<void accept(BigInteger amount, PayableContract buyer, Offer offer)>, that is,
the two type variables \<S> and \<O> have been \emph{erased} and replaced with their
respective upper bound, as illustrated in Fig.\,\ref{figure.java_generics_erasure}.

\begin{figure}[ht]
\centering
\includegraphics[width=.9\linewidth]{figures/java_generics_erasure}
\caption{Example of generics implementation by erasure in Java.}
\label{figure.java_generics_erasure}
\end{figure}

### 2022-02-21 cluster_computing/017_takamaka.tex
Edited the shared entity description
figures

This section gives a short introduction to the Takamaka subset of
Java that we use for writing smart contracts in this paper.
This language has been introduced in~\cite{Spoto19}. A full tutorial
is available online, as part of the documentation of the Hotmoka
blockchain that runs smart contracts written
in Takamaka~\cite{hotmoka_repository}. This section introduces only
the essential notions that are needed to understand the subsequent sections.
The hierarchy of the classes described in this section is in Fig.~\ref{fig:hierarchy-entities}.
In the following we will report a simplified presentation of the code of some
of such classes. The full code is in the Github repository of Hotmoka.

@@

Method calls started from outside the blockchain (for instance, from a client such as a wallet
or from a web application), must specify an already existing \<ExternallyOwnedAccount>
as caller. This account will pay for the gas of the execution. The blockchain will accept
the call only if is signed with the private key that matches the public key provided to
the constructor of the account when it was created. Method \<publicKey> allows one to
recover that public key and method \<nonce> allows one to get a progressive identifier that
can be used to distinguish successive calls with the same account, to force their order
of execution and to avoid replaying.
All that is very similar to Solidity, except for the fact that externally owned accounts
are actual Java objects inside the blockchain, not just an abstraction of a public key.

Neither the Takamaka language nor the Hotmoka blockchain dictate a specific consensus mechanism.
Both proof of work and proof of stacke can be used, for instance. In particular, if proof of stake
is used, then each validator node of the blockchain must specify a
\<io.takamaka.code.governance.Validator> object, that plays the role of the
banking account where the validation rewards of the node get accumulated (see Fig.~\ref{fig:validator}).
It is a special externally owned account, with an extra \<id> method that provides the
identifier of the validator node inside the blockchain network. This identifier depends from the
specific blockchain implementation.
For instance, the subclass
\<io.takamaka.code.governance.tendermint.TendermintED25519Validator>
implements \<id> as for the Tendermint blockchain engine, that is,
as the first $40$ hexadecimal digits of the sha256 digest of the Base64-encoded public key.

\begin{figure}[t]
  \begin{center}
    \begin{lstlisting}[language=Takamaka]
public class Validator extends ExternallyOwnedAccount {

@@

>>>>

This section gives a short introduction to the Takamaka subset of
Java that we use for writing smart contracts in this paper.
This language has been introduced in~\cite{Spoto19}. A full tutorial
is available online, as part of the documentation of the Hotmoka
blockchain that runs smart contracts written
in Takamaka~\cite{hotmoka_repository}. This section introduces only
the essential notions that are needed to understand the subsequent sections.
The hierarchy of the classes described in this section is in Fig.~\ref{fig:hierarchy-entities}.
In the following we will report a simplified presentation of the code of some
of such classes. The full code is in the Github repository of Hotmoka \footnote{\url{https://github.com/Hotmoka}}.

@@

\begin{figure}[th]
  \begin{center}
    \begin{lstlisting}[language=Takamaka]
public class Validator extends ExternallyOwnedAccount {

@@

Method calls started from outside the blockchain (for instance, from a client such as a wallet
or from a web application), must specify an already existing \<ExternallyOwnedAccount>
as caller. This account will pay for the gas of the execution. The blockchain will accept
the call only if is signed with the private key that matches the public key provided to
the constructor of the account when it was created. Method \<publicKey> allows one to
recover that public key and method \<nonce> allows one to get a progressive identifier that
can be used to distinguish successive calls with the same account, to force their order
of execution and to avoid replaying.
All that is very similar to Solidity, except for the fact that externally owned accounts
are actual Java objects inside the blockchain, not just an abstraction of a public key. An exemplification of a call made to a \<PayableContract> is reported in Fig.\,\ref{figure.takamaka_payable_contract}

\begin{figure}[ht]
\centering
\includegraphics[width=.9\linewidth]{figures/takamaka_payable_contract}
\caption{}
\label{figure.takamaka_payable_contract}
\end{figure}

Neither the Takamaka language nor the Hotmoka blockchain dictate a specific consensus mechanism.
Both proof of work and proof of stake can be used, for instance. In particular, if proof of stake
is used, then each validator node of the blockchain must specify a
\<io.takamaka.code.governance.Validator> object, that plays the role of the
banking account where the validation rewards of the node get accumulated (see Fig.~\ref{fig:validator}).
It is a special externally owned account, with an extra \<id> method that provides the
identifier of the validator node inside the blockchain network. This identifier depends from the
specific network.
For instance, the subclass
\<io.takamaka.code.governance.tendermint.TendermintED25519Validator>
implements \<id> as for the Tendermint blockchain engine, that is,
as the first $40$ hexadecimal digits of the sha256 digest of the Base64-encoded public key.
